Bayesian exponential random graph modeling of KWL Burial network

### 2020-06-03 analysis/paper/paper.Rmd
udpate method section

We use burial data collected from published excavation report, and original excavation records with detailed stratigraphy and context description. Given that burials in Chinese phase are far less than previous two phases, we only describe our observations for those burials for comparison. Burial data provides information about direction, size, depth, age, gender, classes and quantity of grave goods, and spatial relationship between them. We consider the effects of both exogenous and endogenous factors to generate meaningful explanation of the characteristics of burial network and tendency of burials correlate with each other. To be specific, our purpose is to understand the underlying mechanisms of the network structures and how the relationship influence on the pattern of accumulation of wealth during different foreign contact periods. We focus on connectivity, clusters, centralization, and homophily among those burials that could reflect the social relations between individuals guided by social structures. Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but how does the degree changes over time and what variables decide the changes remain unclear. To answer the research question for the shifts in social inequality to a more wealth differentiation during and after the arrival of the foreign presences, we hypothesize:

• Hypothesis 1: There are significant reciprocal structure effects in burial network before the contact
• Hypothesis 2: There are significant transitivity structure effects in burial network after the contact
• Hypothesis 3: There are significant popularity structure effects in in burial network after the contact

The computational processes in this paper were performed by the ergm package for simulation and model specification and then bergm package using R programming [@Hunter2008; @Morris2008; @Caimo2014]. We set the number of chains to eight. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10000. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with a set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.

>>>>

We use burial data collected from the published excavation report, and original fieldwork note with detailed descriptions about context and stratigraphy. Given that burials in Chinese phase are less than previous two phases, we only compare these burials by their grave goods and consitions without further network analysis. Burial data provides information about the location, orientation, size, depth, age, gender, quantity of grave goods, and spatial relationship among them. We consider the effects of both exogenous and endogenous factors for meaningful explanation of the characteristics of burial network and tendency of burials correlate with each other. To be specific, our purpose is to understand the underlying mechanisms of the network structures and how their relationships have an influence on the accumulation of wealth between different foreign contact periods. We focus on connectivity, clusters, centralization, and homophily those properties of network that can reflect the social relations between individuals guided by social structures. Previous studies have shown that there is evidence of economic inequality between Kiwualn burials. However, the degree of inequality over time, and which variable decides the changes remain unclear.

We hypothesize that after the arrival of the European presence in the 17th century, a group of ambitious individuals competed with each other for trading with the Europeans due to the high value of trade goods in their society, which led to the accumulation of high quality trade goods. Social inequality gradually increased among Kiwulan society, which can be observed from the burials from a more equal distribution of grave goods to a more differentiation in goods accumulation. We hypothesize that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects in burial network after the European presence.

We performed network analysis with Bayesian inference with the ergm package for simulation and model specification [@Hunter2008; @Morris2008], and then bergm package for Bayesian modeling using R programming [@Caimo2014]. We set the number of chains to eight. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10000. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with a set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
